---
title: "PSTAT174 LAB03"
author: "Yanru Fang"
date: "10/18/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

(a) Express the model in terms of the back shift operator, B.

$(1-0.8B+0.12B^2)X_t = Z_t$

(b) Determine whether the model is causal and/or invertible. (Hint: use polyroot()).\
The autoregressive models are by nature invertible. The AR polynomial for this process has the
factorization $\phi(z) = 1-0.8z+0.12z^2$ and $z=5$ or $z=\frac{5}{3}$. Since these zeroes lie outside the unit circle, we conclude that $X_t$ is a causal AR(2).\
```{r echo=TRUE}
polyroot(c(1,-0.8,0.12))
```

(c) We simulate 200 observations from this AR(2) model with the following code:
```{r echo=TRUE}
set.seed(1234)
ar2 <- arima.sim(model = list(ar = c(0.8,-0.12),sd = 1),n = 200)
```
```{r echo=TRUE}
theo_acf <- ARMAacf(ar = c(0.8,-0.12),lag.max = 20, pacf = FALSE)
op <- par(mfrow = c(1,2))
plot(theo_acf,type = "h",ylim = c(-1,1),
main = "Theoretical ACF",
col = "red",
ylab = expression(rho[X](h)), xlab = "h")
abline(h = 0) 
acf(ar2,lag.max = 20,
main = "Sample ACF",
ylim = c(-1,1),
xlab = "h",
ylab = expression(hat(rho)[X](h)))
```
```{r echo=TRUE}
theo_pacf <- ARMAacf(ar = c(0.8,-0.12),lag.max = 20, pacf = TRUE)
op <- par(mfrow = c(1,2))
plot(theo_pacf,type = "h",ylim = c(-1,1),
main = "Theoretical PACF",
col = "red",
ylab = expression(rho[X](h)), xlab = "h")
abline(h = 0)
pacf(ar2,lag.max = 20,
main = "Sample PACF",
ylim = c(-1,1),
xlab = "h",
ylab = expression(hat(rho)[X](h)))
```
(d) Compare ACF and PACF plots in (c).\
Compared with ACFs, AR models are characterized by PACF=0 after lag k. However, ACFs do not become zeros. In the AR(2) model, theoretical PACFs cut off after lag k=2.
