---
title: "PSTAT174 Lab05"
author: "Yanru Fang"
date: "11/2/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r echo=TRUE}
library(tsdl)
milk <- subset(tsdl, 12, "Agriculture")[[3]]
plot(milk)
```

(a). Explain why the series milk looks not stationary. To make series milk stationary, please difference at lag 12 and then at lag 1.
```{r echo=TRUE}
acf(milk, lag.max = 60, main = "")
title("ACF: Original Time Series", line = -1, outer=TRUE)
```
From the plot of the milk, there is clear non-stationary and trend. To make series milk stationary, we take difference at lag 12 and then lag 1.\

```{r echo=TRUE}
y_1 = diff(milk, 12)
acf( y_1, lag.max = 60, main = "")
title("ACF: First Differencing of Time Series", line = -1, outer = TRUE)
```

```{r echo=TRUE}
y_2 = diff(milk, 1)
acf( y_2, lag.max = 60, main = "")
title("ACF: Second Differencing of Time Series", line = -1, outer = TRUE)
```

(b). Let $Y_t$ be the series ddmilk, that is, $Y_t = (1-B)(1-B^{12})X_t$. Plot the ACF and PACF of $Y_t$ with lagmax equal to 50 and with lagmax equal to 12.
```{r echo=TRUE}
op <- par(mfrow = c(1,2))
y_12 = diff(y_2, 12)
acf( y_12, lag.max = 50, main = "ACF: First and Seasonally Differenced Time Series")
pacf( y_12, lag.max = 12, main = "PACF: First and Seasonally Differenced Time Series")
```

(c). Now, we assume that $Y_t$ corresponds to a SARIMA model. Determine possible candidate models SARIMA for the series $Y_t$.
The ACF shows a strong peak at h = 0s and smaller peaks appearing at h = 1s, 3s.\
The PACF shows two strong peaks at h = 1s and smaller peaks at h = 0.4s, 0.6s.\
The PACF cuts off at lag h=0.6 or 1.\

(d). Choose one stationary and invertible model for this data set, and write down your fitted model.
$SARIMA(p=1, d=1, q=0)(P=1, D=1, Q=1)_{s=12}$
$SARIMA(p=0.6, d=1, q=0)(P=3, D=1, Q=1)_{s=12}$


