---
title: "PSTAT174 Lab02"
author: "Yanru Fang"
date: "10/11/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
library("readxl")
motor_death_month = read_excel("Deaths by Month Table.xlsx",
                               skip=1, n_max=391)[-c(1,2),]
#delete summery information
delete = c(1:30)*13
motor_death_month = motor_death_month[-delete,]
#reorder the data.
groups = split(motor_death_month, rep(1:30,each=12))
motor_death_month = do.call(rbind, rev(groups))
```
## Question 1:
(a). Choose the Deaths column and convert the data to ts format.
```{r echo=TRUE}
deaths <- motor_death_month$Deaths
deaths
deaths_ts <- ts(deaths, start = c(1992, 1), frequency = 12)
deaths_ts
```

(b). Plot the data versus each month.
```{r echo=TRUE}
death_per_vehicle = ts(motor_death_month[,4], start = c(1992,1), 
                       frequency = 12)
plot(death_per_vehicle, main = "Time Series from Year 1992 to 2021")
```

(c). Plot the data versus index number from 1, 2, . . . n.
```{r echo=TRUE}
plot(1:length(death_per_vehicle),death_per_vehicle, main =
"Time Series from Year 1992 to 2021", type = 'l',xlab='index')
index = 1: length(death_per_vehicle)
trend <- lm(death_per_vehicle ~ index)
abline(trend, col="red")
abline(h=mean(death_per_vehicle) , col='blue')
```

(d). Please cut the start of the data and the end, e.g. your data will start at Jan 2009 and
end at Dec 2018 and use the last three year as testing data.
```{r echo=TRUE}
select <- c(1:(17*12))
data_split <- death_per_vehicle[-c(select)]
training = data_split[1:(10*12)]
testing = data_split[((10*12)+1):(13*12)]
```

(e). Create quarterly average. Compare roughness/smoothness of the data.
```{r echo=TRUE}
library(zoo)
smooth_data_5m = zoo::rollmean(death_per_vehicle, k = 5, fill = NA)
smooth_data_12m = zoo::rollmean(death_per_vehicle, k = 12, fill = NA)
plot(death_per_vehicle, main = "Time Series from Year 1992 to 2021")
lines(smooth_data_5m, col='blue' )
lines(smooth_data_12m, col='red')
legend(2013, 2, legend=c("Original data","5-month average", "12-month average"),
col=c("black" ,"blue", "red"), lty=1, cex=0.8)
```
From the plot, we can get the roughness or smoothness of this data. Compared with the 5-month data, the longer interval 12-month data is more smoother.

## Question 2:
```{r echo=TRUE}
#1. set the random seed
set.seed(444) #any integer number
z_t <- rnorm(200,0,1)
plot(z_t,xlab = "t",ylab = expression(z[t]),type = "l",main = "White Noise")
```

(a). Plot Zt and Xt together with different colors. Do you observe any difference?
```{r echo=TRUE}
x_t = filter(z_t, filter = rep(1/5,5), sides = 2, method = "convolution")
plot(z_t,xlab = "t",ylab = expression(Z[t]),type = "l",
main = "Smoothing Process and White Noise")
# Plot of Smoothing Process
lines(x_t,col = "blue")
legend("topright",c("WN", "Smoothing Process"),col = c("black","blue"),lty = 1)
```
I saw the smoothness of those two graphs are different. The graph of X is smoother.

(b). Plot sample acfs of Zt and Xt and compare the two plots.
```{r echo=TRUE}
acf(z_t, main="ACF of White Noise")
acf(x_t,na.action = na.pass, main="ACF of Smoothing Process")
```

